---
title: "Attempting finality at individual level and group/level MHF COVID-19 Analyses"
author: "Zachary Levine"
date: "`r Sys.Date()`"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Individual level variables of interest
1. Met minutes
1. Age

##Group level variables of interest.
1. Volume of visits
1. Attendance
1. %Females

We'll be using the automatic ARIMA model fitting machinery from the *forecast* package to simplify this a bit.

```{r}
library(lubridate)
library(forecast)
library(rstatix)
```

First let's generate the three data frames of interest; mets, prescheduled appointments (volume), and noshows. Split each into only appointments in 2019 and only appointments in 2020.
```{r}

##Change the year to select the year
splitbyyear <- function(df, year = "2019"){
  df[lubridate::year(df$Date) == year,]
}
mets2019 <- splitbyyear(RMHF::read_individual_data(fn = "mets.csv"), "2019")
mets2020 <- splitbyyear(RMHF::read_individual_data(fn = "mets.csv"), "2020")
##Order by date.
mets2019 <- mets2019[order(mets2019$Date),]
mets2020 <- mets2020[order(mets2020$Date),]

volume2019 <- splitbyyear(RMHF::read_individual_data(fn = "volume.csv"), "2019")
volume2020 <- splitbyyear(RMHF::read_individual_data(fn = "volume.csv"), "2020")
##Order by date.
volume2019 <- volume2019[order(volume2019$Date),]
volume2020 <- volume2020[order(volume2020$Date),]
```

Now, let's split off into seperate analyses for each variable. For each, we ask two questions:
1. Does 2019 differ from 2019?
1. If that difference exists, is it before and after the implementation of COVID-19 social distancing in Ontario?

####  Individual level data - Met minutes, sex and age.

First, aggregate data on the daily level.
```{r}
##Should do the heavy lifting for us.
aggregate_date <- function(date, df, colselect = "Mets"){
  dfdate <- df[df$Date == date, colselect]
  return(mean(dfdate))
}
##Vectorized, gives data at the daily level.
meanmets2020 <- sapply(unique(mets2020$Date), aggregate_date, df = mets2020)
meanmets2019 <- sapply(unique(mets2019$Date), aggregate_date, df = mets2019)
meanage2020 <- sapply(unique(volume2020$Date), aggregate_date, df = volume2020, colselect = "Age")
meanage2019 <- sapply(unique(volume2019$Date), aggregate_date, df = volume2019, colselect = "Age")
```

We'd expect some autoregressive (ARIMA) structure in the data, so let's check for that.

```{r}
library(forecast)
auto.arima(meanmets2019)
auto.arima(meanmets2020)
auto.arima(meanage2019)
auto.arima(meanage2020)
```

ARIMA models mostly come up up as white noise processes. So we can't predict anything. Defaulting to anovas and t - tests.

##### Mets

```{r}
##Normal stuff, t - test between data in different years.
library(rstatix)
t.test(meanmets2020, meanmets2019[1:length(meanmets2020)], paired = TRUE)
covidIndex <- as.numeric(min(which(unique(mets2020)$Date == lubridate::ymd("2020/03/16"))))
df2020before <- mets2020[1:covidIndex]
df2020after <- mets2020[covidIndex + 1:length(df2020)]
```

####  Age

```{r}
##Normal stuff, t - test between data in different years.
t.test(meanage2020, meanage2019[1:length(meanage2020)], paired = TRUE)
covidIndex <- as.numeric(min(which(unique(mets2020)$Date == lubridate::ymd("2020/03/16"))))
df2020before <- meanage2020[1:covidIndex]
df2020after <- meanage2020[covidIndex + 1:length(df2020)]
```

So there were no individual level differences between years. Or due to COVID. On to group - level stuff.

##Group level variables of interest.
1. Volume of visits
1. Attendance
1. %Females

####  Volume of visits.

```{r}
df <- RMHF::read_group_data()
df2019 <- as.numeric(df[df$Year == "2019",]$"Prescheduled appointments")
df2020 <- as.numeric(df[df$Year == "2020",]$"Prescheduled appointments")

##Between 2019 and 2020.
t.test(df2019, df2020, paired = TRUE)
##Withint 2020 before and after COVID-19 closure.
t.test(df2020before, df2020after, paired = TRUE)
```



####  Attendance

```{r}
df <- RMHF::read_group_data()
df2019 <- as.numeric(df[df$Year == "2019",]$"% of patients who were no-shows")
df2020 <- as.numeric(df[df$Year == "2020",]$"% of patients who were no-shows")

##Between 2019 and 2020.
t.test(df2019, df2020, paired = TRUE)
##Withint 2020 before and after COVID-19 closure.
t.test(df2020[1:5], df2020[6:10], paired = TRUE)
```


####  %Females

```{r}
df <- RMHF::read_group_data()
df2019 <- as.numeric(df[df$Year == "2019",]$"% Females")
df2020 <- as.numeric(df[df$Year == "2020",]$"% Females")


##Between 2019 and 2020.
t.test(df2019, df2020, paired = TRUE)
##Withint 2020 before and after COVID-19 closure.
t.test(df2020[1:5], df2020[6:10], paired = TRUE)
```


## Correlations between attendance and met-minutes.

```{r}
mets2020 <- RMHF::read_group_data()[RMHF::read_group_data()$Year == "2020",]$"Met-minutes"
mets2019 <- RMHF::read_group_data()[RMHF::read_group_data()$Year == "2019",]$"Met-minutes"
att2020 <- RMHF::read_group_data()[RMHF::read_group_data()$Year == "2020",]$"% of patients who were no-shows"
att2019 <- RMHF::read_group_data()[RMHF::read_group_data()$Year == "2019",]$"% of patients who were no-shows"

#The ith element in meanmets is the mean of the mets collected on ith day in that year.
cor(as.numeric(mets2019), as.numeric(att2019))
cor(as.numeric(mets2020), as.numeric(att2020))
```
